---
title: "riverside_mapping"
author: "Joyce Nguy"
date: "2023-05-18"
output: html_document
---


##Load Packages

```{r cars}
library(sp)
library(rgdal)
library(tigris)
```
#Load in Census Data (all data can be found here: https://ucla.box.com/s/3rms8zmpcgfbnvlndapziiwe3m06zlm4)

```{r}
census_2020 <-read_csv("/Users/student1/Downloads/census_2020.csv")
```

#cleaning data 

```{r}
library(tidyverse)
library(dplyr)
latino_cvap <- cvap_2020%>% 
  mutate(hispanic_cvap = Hsp_CIT_19) %>% 
  dplyr::select(BLOCK20, hispanic_cvap)

#census_2020$Geo_FIPS <- paste0("0", census_2020 $Geo_FIPS)  

census_2020 <- census_2020 %>% 
  mutate(hisp = SE_T028_001, 
         total_pop = SE_T028_000)

latino_census <- census_2020 %>%  
  dplyr::select(Geo_FIPS, Geo_CT, hisp, total_pop) %>% 
  mutate(percent_hispanic = (hisp/total_pop)*100)

```


## Getting Census Tract Boundary Data using Tigris

```{r pressure, echo=FALSE}

# Call tracts(): riverside_blocks
riverside_tracts <- tracts(state = "CA", county = "Riverside")

# Plot 
plot(riverside_tracts)

str(riverside_tracts)


```
#Merge tracts with census data
```{r}
typeof(latino_census$Geo_CT)
typeof(riverside_tracts$TRACTCE)
options(scipen = 999)

#latino_census$Geo_FIPS <- paste0("0", latino_census$Geo_FIPS)

any(duplicated(latino_census$Geo_CT))
any(duplicated(riverside_tracts$TRACTCE))

all(riverside_tracts$TRACTCE %in% latino_census$Geo_CT)
all(latino_census$Geo_CT %in% riverside_tracts$TRACTCE)
riverside_tracts_merge <- merge(riverside_tracts, latino_census, by.x="TRACTCE", by.y = "Geo_CT")

map_data <- merge(riverside_tracts, latino_census, by.x = "TRACTCE", by.y = "Geo_CT", all.x = TRUE)
```

#Regular Chloropleth Map
```{r}
# Plot the chloropleth map
ggplot() +
  geom_sf(data = map_data, aes(fill = percent_hispanic)) +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Percentage of Hispanics in Riverside County",
       fill = "Percent Hispanic") +
  theme_minimal()
```

#overlay
```{r}
#Mapping the Supervisory Districts

library(raster)
supervisory_district<-shapefile("Supervisorial_Districts_2021")
plot(supervisory_district)

library(RColorBrewer)
library(sf)

#make this an sf object
supervisory_district <- st_as_sf(supervisory_district)
```


```{r}
#Check CRS
st_crs(map_data)
st_crs(supervisory_district)

# Transform CRS if needed
supervisory_district <- st_transform(supervisory_district, crs = st_crs(map_data))

#map
map <- ggplot() +
  geom_sf(data = riverside_tracts_merge, aes(fill = percent_hispanic), color = "grey") +
  geom_sf(data = supervisory_district, fill = NA, color = "black", size = 3) +
  coord_sf() +
  labs(title = "Percentage of Latinos in Riverside County",
       subtitle = "Census Tracts",
       fill = "Percent Latino") +
  scale_fill_gradientn(colors = brewer.pal(9, "YlOrRd"),
                       breaks = c(25, 50, 75, 100),
                       labels = c(25, 50, 75, 100),
                       limits = c(0, 100)) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.background = element_rect(color = "black", fill = "white"),
        legend.box.background = element_rect(color = "black", fill = "white"),
        legend.key = element_rect(color = "black", fill = "white"))

# Display the map
map

ggsave("riverside_latino_map.png", plot = map, width = 10, height = 8, dpi = 300)

```


#CVAP MAP 

```{r}
#Load in Census Data 
cvap_2020 <- read_csv("/Users/student1/Downloads/Riverside Mapping/cvap_2020.csv")
names(cvap_2020)
```

#Clean Data

```{r}
library(dplyr)
#filter for riverside county by tract
filtered_data <- cvap_2020 %>%
  filter(str_detect(geoname, "Riverside County"))

# View the filtered data
View(filtered_data)

#filter for hispanic and latino, create latino percentage
filtered_data <- filtered_data %>%
  filter(lntitle %in% c("Hispanic or Latino", "Total")) %>%
  group_by(geoid) %>%
  mutate(percentage_hispanic = cvap_est[lntitle == "Hispanic or Latino"] / cvap_est[lntitle == "Total"] * 100) #%>%
#  filter(!is.na(percentage_hispanic))
#2 NA Tracts 

# View the resulting data
View(filtered_data)

#distinct geoid
filtered_data <- filtered_data %>%
  distinct(geoid, .keep_all = TRUE)

##make a tract column with last 6 digits 
filtered_data$tract <- substr(filtered_data$geoid, nchar(filtered_data$geoid) - 5, nchar(filtered_data$geoid))

# View the resulting data
View(filtered_data)
```


##MAP the CVAP 2020 data 

```{r}
any(duplicated(filtered_data$tract))
any(duplicated(riverside_tracts$TRACTCE))

all(riverside_tracts$TRACTCE %in%filtered_data$tract )
all(filtered_data$tract %in% riverside_tracts$TRACTCE)

cvap_map_data <- merge(riverside_tracts, filtered_data, by.x = "TRACTCE", by.y = "tract", all.x = TRUE)

# Transform CRS if needed
supervisory_district <- st_transform(supervisory_district, crs = st_crs(cvap_map_data))

#map
cvap_map <- ggplot() +
  geom_sf(data = cvap_map_data, aes(fill = percentage_hispanic), color = "grey") +
  geom_sf(data = supervisory_district, fill = NA, color = "black", size = 3) +
  coord_sf() +
  labs(title = "Percentage of Latinos in Riverside County - Citizen Voting Age Population",
       subtitle = "Census Tracts",
       fill = "Percent Latino") +
  scale_fill_gradientn(colors = brewer.pal(9, "YlOrRd"),
                       breaks = c(25, 50, 75, 100),
                       labels = c(25, 50, 75, 100),
                       limits = c(0, 100)) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.background = element_rect(color = "black", fill = "white"),
        legend.box.background = element_rect(color = "black", fill = "white"),
        legend.key = element_rect(color = "black", fill = "white"))

# Display the map
cvap_map

ggsave("riverside_latino_cvap_map.png", plot = cvap_map, width = 10, height = 8, dpi = 300)

```


Tmap
```{r}
library(tmap)

# Convert the merged data to a tm object
tm_data <- tm_shape(riverside_tracts_merge) +
  tm_polygons(col = "percent_hispanic", palette = "Blues", title = "Percent Hispanic") +
  tm_layout(main.title = "Percentage of Hispanics in Riverside County",
            legend.outside = TRUE,
            legend.position = c("right", "bottom"))

# Customize the legend box
tm_data <- tm_data +
  tm_legend(outside = TRUE, 
            bg.color = "white",
            frame = TRUE,
            legend.outside.size = 0.5,
            title.size = 0.9,
            title.bg.color = "white")

# Display the map
tm_view(tm_data)

#save
tmap_save(tm_data, filename = "riverside_map.png")

```




## Tidycensus? 

```{r}
# look up variables 
variables <- load_variables(year = 2020, dataset = "pl")
acs_variables <- load_variables(year = 2015, dataset = "acs5")
library(tidycensus)
riverside_race <- get_decennial(
  geography = "tract",
  state = "CA",
  county = "Riverside",
  variables = c(Hispanic = "P2_002N"),
  summary_var = "P2_001N",
  year = 2020,
  geometry = TRUE
) %>%
  mutate(percent = 100 * (value / summary_value))
```


```{r}
library(tmap)
base_map <- tm_shape(riverside_race) + 
  tm_polygons(col = "percent",
          style = "quantile",
          n = 5,
          palette = "Purples",
          title = "2020 US Census") + 
  tm_layout(title = "Pct Latino in Census Tract",
            frame = FALSE,
            legend.outside = TRUE)

str(base_map)
```




##Riverside Mapping by Block 

```{r}
#create riverside blocks 
riverside_blocks <- blocks(state = "CA", county = "Riverside")

# Plot 
plot(riverside_blocks)

str(riverside_blocks)
```

#Load in the Block Data from Statewide District - Adjusted
```{r}
adjusted_data <- read_csv("/Users/student1/Downloads/Riverside Mapping/county_065_PL94_2020_Adjusted_P24_DOJ_Block_csv/riverside_block_adjusted.csv")

#create percentage hispanic
adjusted_data <- adjusted_data %>%
  mutate(percentage_hispanic = (`Hispanic Origin` / `Population P2`) * 100)

names(adjusted_data)


any(duplicated(adjusted_data$BLOCK20))
any(duplicated(riverside_blocks$GEOID20))

all(riverside_blocks$GEOID20 %in%adjusted_data$BLOCK20 )
all(adjusted_data$BLOCK20  %in% riverside_blocks$GEOID20)

adjusted_map_data <- merge(riverside_blocks, adjusted_data, by.x = "GEOID20", by.y = "BLOCK20", all.x = TRUE)

# Transform CRS if needed
supervisory_district <- st_transform(supervisory_district, crs = st_crs(adjusted_map_data))

#map
adjusted_map <- ggplot() +
  geom_sf(data = adjusted_map_data, aes(fill = percentage_hispanic), color = "grey") +
  geom_sf(data = supervisory_district, fill = NA, color = "black", size = 3) +
  coord_sf() +
  labs(title = "Percentage of Latinos in Riverside County",
       subtitle = "By Census Block",
       fill = "Percent Latino") +
  scale_fill_gradientn(colors = brewer.pal(9, "YlOrRd"),
                       breaks = c(25, 50, 75, 100),
                       labels = c(25, 50, 75, 100),
                       limits = c(0, 100), 
                       na.value = "white") +  # Set NaN values to white+
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.background = element_rect(color = "black", fill = "white"),
        legend.box.background = element_rect(color = "black", fill = "white"),
        legend.key = element_rect(color = "black", fill = "white"))

# Display the map

sum(is.na(adjusted_data$percentage_hispanic))
sum(is.na(adjusted_data$percentage_hispanic) | is.nan(adjusted_data$percentage_hispanic))

ggsave("riverside_adjusted_map.png", plot = adjusted_map, width = 10, height = 8, dpi = 300)
```

#Map Raw

```{r}
adjusted_map <- ggplot() +
  geom_sf(data = adjusted_map_data, aes(fill = `Hispanic Origin`), color = "grey") +
  geom_sf(data = supervisory_district, fill = NA, color = "black", size = 3) +
  coord_sf() +
  labs(title = "Number of Hispanics in Riverside County",
       subtitle = "By Census Block") +
  scale_fill_gradientn(colors = brewer.pal(9, "YlOrRd"),
                       breaks = c(0, 250, 500, 750, 1000, 1250, 1500, 1750, 2000),
                       limits = c(0, 2050)) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.background = element_rect(color = "black", fill = "white"),
        legend.box.background = element_rect(color = "black", fill = "white"),
        legend.key = element_rect(color = "black", fill = "white"))

ggsave("riverside_adjusted_map_raw_count.png", plot = adjusted_map, width = 10, height = 8, dpi = 300)

```

#create dot density map 

```{r}
# Convert data to SpatialPointsDataFrame
adjusted_points <- adjusted_map_data %>%
  st_as_sf() %>%
  st_centroid() %>%
  st_coordinates() %>%
  as.data.frame()

# Rename coordinates columns
names(adjusted_points) <- c("x", "y")

# Combine adjusted_map_data with adjusted_points
adjusted_map_data <- bind_cols(adjusted_map_data, adjusted_points)

# Create dot density map
# Adjust colors based on count of Latinos
# Adjust colors based on count of Latinos
dot_map <- ggplot() +
  geom_sf(data = adjusted_map_data, fill = NA, color = "grey") +
  geom_sf(data = supervisory_district, fill = NA, color = "black", size = 3) +
  geom_point(data = adjusted_map_data, aes(x = x, y = y, color = `Hispanic Origin`), size = 0.5, alpha = adjusted_map_data$`Hispanic Origin`/max(adjusted_map_data$`Hispanic Origin`)) +
  coord_sf() +
  labs(title = "Number of Hispanics in Riverside County",
       subtitle = "By Census Block") +
  scale_color_gradient(low = "lavender", high = "purple") +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "none")

dot_map



```
#Remove Census Boundaries

```{r}
adjusted_map_no_bound <- ggplot() +
  geom_sf(data = adjusted_map_data, aes(fill = percentage_hispanic), color = "transparent") +
  geom_sf(data = supervisory_district, fill = NA, color = "black", size = 3) +
  coord_sf() +
  labs(title = "Percentage of Latinos in Riverside County",
       subtitle = "By Census Block",
       fill = "Percent Latino") +
  scale_fill_gradientn(colors = brewer.pal(9, "YlOrRd"),
                       breaks = c(25, 50, 75, 100),
                       labels = c(25, 50, 75, 100),
                       limits = c(0, 100), 
                       na.value = "white") +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.background = element_rect(color = "black", fill = "white"),
        legend.box.background = element_rect(color = "black", fill = "white"),
        legend.key = element_rect(color = "black", fill = "white"))

adjusted_map_no_bound

ggsave("adjusted_map_no_boundt.png", plot = adjusted_map_no_bound, width = 10, height = 8, dpi = 300)

```

